---
title: "Writing Functions in R"
author: "Munguldei Batsaikhan"
date: "December 4, 2023"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## 1. What is a Function?

A function in R is an object containing multiple interrelated statements that are run together in a predefined order every time the function is called. Functions in R can be built-in or user-defined. The main purpose of creating a user-defined function is to optimize a program, avoid repetition of code and make the code more readable.

## 2. Built-in Functions

* Numeric Manipulation: abs, sqrt, ceiling, floor
* Basic stats: mean, median, sd, cor, quantile
* Probability functions:
  * Density functions (ddist): dnorm, dbinom, dpois, dunif
  * Cumulative distribution functions (pdist): pnorm, pbinom, ppois, punif
  * Quantile functions (qdist): qnorm, qbinom, qpois, qunif
  * Random numbers (rdist) : rnorm, rbinom, rpois, runif

## 3. Creating Your Own Functions

To make a user-defined function in R, follow this structure,

```{r, eval=F, echo=T}
function_name <- function(parameters){
  function body 
}
```


## 4. Function Examples

Starting off simple with no parameters,
```{r}
done<-function(){
  print("Yaay, you finished a task!")
}
```
 
Calling your very first function!
```{r}
done()
```


Now, let's include parameters for a more interesting function,
```{r}
#zazz it up with parameters
encourage<-function(n=1){
  quotes<-read.csv("~/stat202/CW/present/quotes.csv")
  selected<-sample_n(quotes, n)
  
  #simple print
  for (i in 1:n) {
    print(selected$quote[i])
  }
}
```

Calling your very first functions with parameters!
```{r}
encourage(2)
```

We can make our function even more sophisticated by adding more parameters,
```{r}
#want something more relevantly inspiring?
encourage2<-function(n=1, key="friendship"){
  quotes<-read.csv("~/stat202/CW/present/quotes.csv")
  
  selected<-c()
  for (j in 1:nrow(quotes)) {
    category<-quotes$category[j]
    if (grepl(key, category)) {
      selected<-append(selected, j)
    }
  }
  
  #if we have desired quotes, random select from found quotes
  if (n > length(selected)) {
      print("Not enough quotes for key theme.")
  }
  selected<-sample(selected, n)
  print(quotes$quote[selected])
}
```

Running improved `encourage`,
```{r}
encourage2()
```

```{r}
encourage2(key="love")
```


## 5. Are Functions Useful?

Absolutely! It saves us the trouble of repetition, ugly code and future debugging issues. Observe that some of the action we performed in R could have been made into functions.

Loading in data set to test function,
```{r}
mb34<-read.csv("~/stat202/CW/midterm/mb34.csv")
mb34<-mb34[,c('SalePrice', 'Bedrooms', 'SquareFeet', 'YearSold', 'YearBuilt')]
mb34$SalePrice<-mb34$SalePrice/1000
lm.price<-lm(SalePrice~Bedrooms+SquareFeet+YearSold+YearBuilt, data=mb34)

```

Function to make model diagnostics easier,
```{r}
#function to check model diagnostics
#pre: df - data frame that contains all predictor variables ONLY included in the lm.obj,
#lm.obj - lm object that was created from df
diagnose<-function(df, lm.obj) {
  df<-na.omit(df)
  par(mfrow=c(2,3))
  #Residual plot for predicted values (for overall linearity)
  plot(lm.obj$fitted.values, lm.obj$residuals, ylab="Residual", xlab="Predicted",
  pch=19, cex.main=0.8, main="Residuals versus Predicted Y")
  abline(h=0, lty=2, lwd=2, col="blue")

  #Residual plot for predictor variables (for each predictor linearity)
  vars = all.vars(formula(lm.obj))[-1]
  for (var in vars) {
    predictor=df[,var]
    if (is.numeric(predictor)) {
      plot(predictor, lm.obj$residuals, ylab="Residual", xlab=var, pch=19, cex.main=0.8, main=paste("Residuals versus", var)) 
      abline(h=0, lty=2, lwd=2, col="blue")
      }
    next
  }
  
  par(mfrow=c(1,2))
  #Histogram of residuals
  hist(lm.obj$residuals, xlab="Residuals", 
       cex.main=0.8, main="Histogram of Residuals")

  #Normal QQ Plot
  qqnorm(lm.obj$residuals)
  qqline(lm.obj$residuals)
}
```

```{r}
diagnose(mb34, lm.price)
```

## 6. Sources
* https://www.kaggle.com/datasets/manann/quotes-500k/
* https://ashki23.github.io/r-functions.html
* https://www.statmethods.net/management/functions.html
* https://www.dataquest.io/blog/write-functions-in-r/
